Privacy Analysis of a Mobile Sensor Application

### 2014-12-30 paper.tex
started subsubsection for alternative definition of privacy

\noindent
Defining privacy is a challenge which seems impossible. This is well put to words by Serge Gutwirth, who notes:
\begin{quote}
The notion of privacy remains out of the grasp of every academic chasing it. Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive. \cite{Gutwirth}
\end{quote}
In this section we introduce the concept of privacy as control over information.
We specify kinds of information we have to deal with in the context of privacy.
And finally we outline the impact mobile sensors can have on one's privacy.

@@

%One particular scholar, who did not follow this way of thinking is Charles Fried.
One possible definition of privacy is given by Charles Fried in his work \textit{An Anatomy of Values}\cite{CFried:Privacy}.
He questioned, why we are intuitively sensitive to privacy violations.
But he did not assert privacy as an intrinsic value by itself, he rather stated:
\begin{quote}
\textbf{Privacy is} not simply an absence of information about us in the minds of others;
rather it is \textbf{the control we have over information about ourselves}. \cite{CFried:Privacy}
\end{quote}
According to Fried privacy is our ability to create and modulate the relationships to other humans.
Which makes us intuitively sensitive for its violation, because this affects what defines our identity.
Also, the ability to build and maintain relationships is essential to society.
This is what makes privacy valuable and deserving of protection.

@@

We share information of great intimacy with persons we deem trustworthy we do not share with anyone else. 
Moreover, we trust those persons to not reveal information about us to others by not constantly monitoring them, i.e. respecting their privacy.
Trust needs the possibility of unknown failure.
If we would constantly monitor our partners, they cannot fail unnoticed nor can they willingly share intimate information with us.
By trusting them as described before, we create the possibility for them to share intimate information with us.

\subsubsection{Relationship Modulation.}
Fried argues, depending on the person we are interacting with, we are changing the degree of intimacy we share.
If we are to share private information with others, we are most likely aware that this action is privacy related.
In that case we are able to selectively disclose information along the two dimensions of quantity and quality.

\subsubsection{Alternative Definitions.}
So privacy and its possibility, according to Fried, is the ability to build and maintain relationships between humans.
Thus it is essential to society, \cite{CFried:Privacy}.
Using his anatomy of privacy as foundation of our analysis is suitable because of two points.
At first, Fried's study on the understanding of privacy provided a great contribution to the research on the same term in philosophy \cite{sep-privacy} and computer science \cite{SociotechnicalArchitectureForOnlinePrivacy}.
Secondly, despite the fact his text was published in 1970, he already included technologies to its viewpoint that did not only monitor location, but also record biometric data.

@@

\begin{LGContent}
In Fried's definition of privacy as control over "information about onself'', the specification of what constitues such information remains open.
There is a vast amount of information that relates to a person and we need to get a better understanding in order to perform a thorough analysis.
To this end we use the the categorization by Friedewald, Finn and Wright \cite{7ToP} called the \textit{Seven Types of Privacy}.
The seven types of privacy are an extension to the four types of privacy by Roger Clarke \cite{RClarke:4ToP}, which are:
\begin{itemize}
\item Privacy of the Person
\item Privacy of Personal Behaviour
\item Privacy of Personal Communication
\item Privacy of Personal Data
\end{itemize}
It is important to note, that this categories do not form no taxonomy, since the categories are not mutually exclusive.
For instance a written email is considered personal communication as well as personal data stored on a computer.

@@

\end{LGContent}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{diagrams/png/implicit-sensor-privacy-matrix.png}

\caption{Live+Gov Implicit Sensor-Privacy Matrix}
\label{figure:Live+Gov Implicit Sensor-Privacy Matrix}
\end{figure}

@@

@@

\subsubsection{Human Activity Recognition.}
The detection and collection of human activities like walking, standing and running, can interfer with the Privacy of the Person, e.g.~since these movement patterns can be indicators for a person's health. Also, trivially, Privacy of Data and Image is violated.

\subsubsection{Service Line Detection.}
The detection and collection of the service line the user currently uses in the public transportation system allows inference of the location of the user, at least when entering and leaving those service lines at e.g.~bus stops.
Implicit privacy violations apply accordingly.

>>>>

\noindent
Defining privacy is a challenge which seems impossible. This is well put to words by Serge Gutwirth, who notes:
\begin{quote}
The notion of privacy remains out of the grasp of every academic chasing it. Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive. \cite{Gutwirth}
\end{quote}
In this section we introduce the concept of privacy as control over information.
We specify types of information we have to deal with in the context of privacy.
And finally we outline the impact mobile sensors can have on one's privacy.

@@

%One particular scholar, who did not follow this way of thinking is Charles Fried.
One possible definition of privacy is given by Charles Fried in his work \textit{An Anatomy of Values}\cite{CFried:Privacy}.
He questioned, why we are intuitively sensitive to privacy violations.
But he did not assert privacy as an intrinsic value by itself, he rather stated:
\begin{quote}
\textbf{Privacy is} not simply an absence of information about us in the minds of others;
rather it is \textbf{the control we have over information about ourselves}. \cite{CFried:Privacy}
\end{quote}
According to Fried privacy is our ability to create and modulate the relationships to other humans.
Which makes us intuitively sensitive for its violation, because this affects what defines our identity.
Also, the ability to build and maintain relationships is essential to society.
This is what makes privacy valuable and deserving of protection. 
\cite{sep-privacy}\cite{CFried:Privacy}

@@

We share information of great intimacy only with persons we deem trustworthy. 
Moreover, we trust those persons to not reveal information about us to others by not constantly monitoring them, i.e. respecting their privacy.
Trust needs the possibility of unknown failure.
If we would constantly monitor our partners, they cannot fail unnoticed nor can they willingly share intimate information with us.
By trusting them, we create the possibility for them to share intimate information with us.
\cite{CFried:Privacy}

\subsubsection{Relationship Modulation.}
Fried argues, depending on the person we are interacting with, we are changing the degree of intimacy we share.
If we are to share private information with others, we are most likely aware that this action is privacy related.
In that case we are able to selectively disclose information along the two dimensions of quantity and quality.
\cite{CFried:Privacy}

\subsubsection{Alternative Definitions.}
Using Fried's anatomy of privacy for this analysis is suitable because of two points.
Despite the fact his text was published in 1970, he already included technologies to its viewpoint that did not only monitor location, but also record biometric data \cite{CFried:Privacy}. 
Secondly, the concept of control over information does not enforce data minimization.
Fried's definition complies with what we call \textit{self data protection}, where persons communicate their data self-determined.
This is the opposite of \textit{system data protection}, where the collection of certain data is prohibited.

@@

%\begin{LGContent}
In Fried's definition of privacy as control over "information about onself'', the specification of what constitues such information remains open.
There is a vast amount of information that relates to a person and we need to get a better understanding in order to perform a thorough analysis.
To this end we use the the categorization by Friedewald, Finn and Wright \cite{7ToP} called the \textit{Seven Types of Privacy}.
The seven types of privacy are an extension to the four types of privacy by Roger Clarke \cite{RClarke:4ToP}, which are:
\textit{Privacy of the Person},
\textit{Privacy of Personal Behaviour},
\textit{Privacy of Personal Communication},
\textit{Privacy of Personal Data}.
It is important to note, that this categories do not form no taxonomy, since the categories are not mutually exclusive.
For instance a written email is considered personal communication as well as personal data stored on a computer.

@@

%\end{LGContent}

@@

@@

\begin{figure}
\newcommand{\tcell}[1]{\parbox[c]{4cm}{\vspace{3mm}#1\vspace{3mm}}}
\newcommand{\thead}[1]{\tcell{\Large\centering\textbf{#1}}}
\newcommand{\tbody}[1]{\tcell{\Large\centering #1}}
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}{|c||c|c|c|c|c|c|c|}
\hline
& \thead{Pricacy of the Person}
& \thead{Pricacy of Behaviour and Action}
& \thead{Pricacy of Communication}
& \thead{Pricacy of Data and Image}
& \thead{Pricacy of Thoughts and Feelings}
& \thead{Pricacy of Location and Space}
& \thead{Pricacy of Association}
\\\hline\hline
\thead{Geo Position}
& \tbody{0.5} % Pricacy of the Person
& \tbody{0.5} % Pricacy of Behaviour and Action
& \tbody{0} % Pricacy of Communication
& \tbody{1} % Pricacy of Data and Image
& \tbody{0} % Pricacy of Thoughts and Feelings
& \tbody{1} % Pricacy of Location and Space
& \tbody{0.5} % Pricacy of Association
\\\hline
\thead{Motion}
& \tbody{1} % Pricacy of the Person
& \tbody{0.5} % Pricacy of Behaviour and Action
& \tbody{0} % Pricacy of Communication
& \tbody{1} % Pricacy of Data and Image
& \tbody{0} % Pricacy of Thoughts and Feelings
& \tbody{0} % Pricacy of Location and Space
& \tbody{0} % Pricacy of Association
\\\hline
\thead{Networking}
& \tbody{0.5} % Pricacy of the Person
& \tbody{0.5} % Pricacy of Behaviour and Action
& \tbody{0} % Pricacy of Communication
& \tbody{1} % Pricacy of Data and Image
& \tbody{0} % Pricacy of Thoughts and Feelings
& \tbody{1} % Pricacy of Location and Space
& \tbody{1} % Pricacy of Association
\\\hline
%\thead{HAR}
%& \tbody{-} % Pricacy of the Person
%& \tbody{-} % Pricacy of Behaviour and Action
%& \tbody{-} % Pricacy of Communication
%& \tbody{-} % Pricacy of Data and Image
%& \tbody{-} % Pricacy of Thoughts and Feelings
%& \tbody{-} % Pricacy of Location and Space
%& \tbody{-} % Pricacy of Association
%\\\hline
%\thead{SLD}
%& \tbody{-} % Pricacy of the Person
%& \tbody{-} % Pricacy of Behaviour and Action
%& \tbody{-} % Pricacy of Communication
%& \tbody{-} % Pricacy of Data and Image
%& \tbody{-} % Pricacy of Thoughts and Feelings
%& \tbody{-} % Pricacy of Location and Space
%& \tbody{-} % Pricacy of Association
%\\\hline
\end{tabular}
}
\newline
\newline
{\scriptsize 
0: No Impact,
0.5: Light Impact,
1: Hard Impact}
\caption{Sensor Data Privacy Impact Matrix}
\label{figure:Sensor Data Privacy Impact Matrix}
\end{figure}

@@

%\subsubsection{Human Activity Recognition.}
%The detection and collection of human activities like walking, standing and running, can interfer with the Privacy of the Person, e.g.~since these movement patterns can be indicators for a person's health. Also, trivially, Privacy of Data and Image is violated.
%
%\subsubsection{Service Line Detection.}
%The detection and collection of the service line the user currently uses in the public transportation system allows inference of the location of the user, at least when entering and leaving those service lines at e.g.~bus stops.
%Implicit privacy violations apply accordingly.
